Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-08 Burzillo_Trounstine_Replication.Rmd
final edits

```
\setlength\parindent{24pt}
# Introduction

|          There is a large degree of variation in public goods spending across local governments. As a result, many scholars have worked to determine what factors may lead to the underprovision of public goods. Research in the past has associated racial diversity or changes in levels of diversity with the under-provision of public goods [@baqir_public_1999; @hopkins_diversity_2009]. However, @trounstine_segregation_2016 argues that it is racial segregation, not diversity in and of itself, that results in the under-provision of public goods. Trounstine’s analysis consists of two main parts. First, she uses election and demographic data from 25 of America’s largest cities between 1990 and 2010 to run a multilevel mixed-effects linear regression with fixed effects for region and year and with random effects for cities in order to show that political polarization increases with segregation. Trounstine measures segregation with Theil's H Index, which measures the degree to which the diversity of a neighborhood differs from the diversity of the entire city. The second main part of Trounstine’s analysis looks at the ability of the Theil’s H Segregation Index to explain a variety of types of public expenditures by city using a sample of 2,637 cities with 13,742 city-year observations. Using linear regressions with fixed effects for cities and robust standard errors clustered by city, Trounstine finds a significant, negative correlation between segregation and public goods spending that is robust to the inclusion of a variety of relevant controls and an alternative specification in which the number of waterways is used as an instrument for segregation.

@@

|          One concern with the original analysis in @trounstine_segregation_2016 is the exclusion of large amounts of missing data, which substantially constrains the sample size in the regression analysis. For example, the regression analysis in the main specification using the the original racial polarization dataset exludes %`r round((obs_rp - 91)/obs_rp * 100, 0)` of the observations, and the regression analysis in the main specification using the original financial segregation dataset excludes %`r round((obs_fin_seg - 13742)/ obs_fin_seg * 100, 0)` of observations. As an extension of @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. Then, I use the multiply imputed datasets to re-estimate the original models, pooling the results to generate final pooled regression coefficients and parameters.

@@

|          In order to understand what segregation is and why it is important as a political concept, it is critical to understand the literature on the effects of diversity at various geographic levels. Studying diversity can be somewhat difficult and confusing because its effects tend to differ by geographic level. On the neighborhood level, a lack of intra-neighborhood racial diversity has been associated with racial intolerance, resentment, and competition between racial groups [@oliver_paradoxes_2010]. Living within non-diverse or segregated neighborhoods has also been associated with holding negative stereotypes and perceptions about out groups [@eric_oliver_intergroup_2003]. As a result, non-diverese, homogeneous neighborhoods have been associated with increased racial tension and political polarization in comparison to integrated, diverse neighborhoods. However, at the city or metropolitan level, the opposite seems to be true: when considering larger geographic areas, diversity, not homogeneity, is correlated with racial tension, competition, prejudice, lower levels of cooperation, and lower spending on public goods [@oliver_paradoxes_2010; @baqir_public_1999; @hopkins_diversity_2009]. As a result, these theories predict that geographic areas most ripe for racial tension and competition will be those that are diverse overall, but have many homogeneous, non-diverse neighborhoods, meaning that while people of different races co-exist within a larger metropolitan area, they live separately within their own neighborhoods. This situation corresponds to @trounstine_segregation_2016's definition of a severely segregated city. The differing effects of diversity at the neighborhood versus the larger geographic level provide an argument for considering segregation as opposed to simply diversity in local politics because measures of segregation (such as the Theils's H Index used by @trounstine_segregation_2016) take into account the patterns of racial mixing at both the neighborhood and larger geographic level that, alltogether, are important in predicting whether a locality will be characterized by interracial cooperation or political polarization [@trounstine_segregation_2016; @oliver_paradoxes_2010; @bharathi_public_2018].

@@

|          In an attempt to better deal with the problem of missing data in @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. While there are a variety of different imputation methods that could have been employed, multiple imputation (such as the multiple multivariate imputations generated by mice) is desirable because instead of inputting a single value such as the mean for missing values, it instead uses the distribution of the available data to estimate multiple potential values for the missing data. As a result, multiple imputation helps to account for the uncertainty inherent in the imputation process and allows for the calculation of standard errors around estimators. Thus, multiple imputation allows the researcher to more accurately assess the uncertainty of the analysis in general.

@@

|          In Figure 2, we can similarly observe the trends for the financial segregation dataset. In this dataset, there is a very high proportion of missing values for most of the variables. For example, there are `r format(sum(is.na(fin_imp$H_citytract_NHW_i)), big.mark = ",")` missing values for the segregation index, which is the main independent variable, or %`r round(sum(is.na(fin_imp$H_citytract_NHW_i))/nrow(fin_imp)*100,0)` of the data. In this dataset, there are also many more observations for which values are missing for multiple variables in comparison to the racial polarization dataset, as can be seen in the figure on the right in Figure 2. In total, a mere %`r .0058*100` of observations are complete for all variables included in the main specification for the financial segregation dataset.

@@

imp_1_rp <- mice(rp_impute, method = "cart", maxit = 5)

@@

imp_fs <- mice(fin_imp, method = meth, maxit = 1)

@@

```{r rp_densityplot, echo = FALSE, fig.cap = "Figure 3 presents the density plots of the imputed data (red) overlaid on the density plot for the pre-existing data (blue) for all those variables with missing values in the racial polarization dataset. For the most part, the imputed data seem to fit the pre-existing data relatively well. The only cases in which there is some cause for concern were for some of the variables with very few missing values in the racial polarization dataset (such as the indicators for region). However, since there are so few of these values missing in the actual datasets, this was not a major concern.", fig.pos="H"}

@@

kable(imp_1_rp_sum_table_vars, 
      "latex", 
      caption = "\\textbf{Racial Polarization in Segregated Cities using the Multiply Imputed Data: Racial Divide with Multigroup Segregation Index}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 203"),
    general = c("This table shows the results of Trounstine's first model for the effect of segregation on political polarization",
  "using the multi-group Theil's H Segregation Index using the new, multiply imputed datasets. The",
  "sign of the coefficient on the main variable of interest, the segregation index, is positive as in",
  "Trounstine (2016), but the result has become statisically insignificant with the 95% confidence",
  "interval containing zero. Thus while these results still suggest that segregation may be",
  "associated with increased political polarization, they confer a lesser degree of certainty than",
  "Trounstine's original analysis. In general, the standard errors on the coefficients have",
  "decreased slightly."),
           general_title = "Table 1:",
  footnote_order = c("symbol", "general"))%>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|          In the second and third model specifications, which use the two-group Theil's H Segregation Index, the results were similar. The main coefficient on segregation index decreased in magnitude, although it remained statistically significant. The other coefficients except for the biracial indicator are insignificant, and standard errors have reduced slightly in these specifications compared to the original analysis in Trounstine (2016). As a visual comparison of the model coefficients, I include coefficient plots for the diversity and segregation index variables as Figure 4. From this plot, it is clear that across all of the models, the coefficient on diversity is positive and insignificant, while the coefficient on the segregaiton index is positive and, in most cases, significant.

@@

|          For the financial segregation dataset, I similarly fit each of our 5 imputed datasets to the original models from @trounstine_segregation_2016 using this data and then pooled the results for each. As a result of the imputation process, I was able to include an additional `r format(73119 - 13742, big.mark = ",")` observations in the analysis, increasing the sample size by %`r round((73119 - 13742)/ 13742*100,0)`. While this is an improvement, I was unable to impute all of the missing values because of the large percentage of missing values in the original dataset, which limited the performance of the algorithm, and due to time and computing constraints for this project.

@@

kable(imp_felm1_sum_table_vars, 
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for all Cities}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 73,119"),
    general = c("This table shows the results of Trounstine's first model for the effect of segregation on per capita direct",
  "public expenditure using the two-group Theil's H Segregation Index using the new, multiply imputed datasets.",
  "The results present a stark contrast to Trounstine's original results: the main coefficient on segregation index",
  "has decreased to essentially zero and has become statistically insignificant.  The coefficient on diversity has",
  "increased in magnitude and become statistically significant."),
  footnote_order = c("symbol", "general"),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|          Because these results differed to such a degree from @trounstine_segregation_2016's original findings, I performed some checks on the data included in each analysis. I discovered that @trounstine_segregation_2016 excluded most cities below a certain population size because of the nature of the paper's listwise deletion. It may be that small cities are more likely to be missing data, and thus they were more likely to be dropped in @trounstine_segregation_2016's analysis. Referring back to @baqir_public_1999, I decided to divide the orginal sample based upon the populaiton cutoff for cities used in that paper, which determined that the populaiton must be above 25,000 people for an area to be considered a city. Splitting the financial segregation dataset in two, one dataset above and one below the 25,000 population cutoff, I then again imputed missing values for each and used the multiply imputed datasets in Trounstine's original models.

|          Figure 5 displays the population distributions used in each of these analyses. The figure makes it clear that the observations included in @trounstine_segregation_2016's original analysis were not representative of the original data sample with the missing values, as most of the cities with lower populations were excluded as a result of the list-wise data deletion. While the dataset generated from the data imputation on the whole sample also has a different distribution for city populaiton, the range of the data seems to be much more representative of the overall data range in comparison to the @trounstine_segregation_2016 sample. For comparison, I also include the distributions of the multiply imputed datasets above and below the 25,000 populaiton cutoff. For cities above the cutoff, the city population distribution is similar to the distribution in @trounstine_segregation_2016.

@@

imp_fs_pop_overbaqir <- mice(fin_imp_pop_over_baqir, method = meth, maxit = 1)

@@

imp_fs_pop_under_baqir <- mice(fin_imp_pop_under_baqir, method = meth, maxit = 1)

@@

pop_hist_data %>%
  ggplot() +
  geom_histogram(aes(x = log_pop, fill = Analysis), binwidth = .5) +
  geom_vline(aes(xintercept = median_log_pop), color = "red") +
  facet_wrap(~Analysis) +
  coord_cartesian(xlim = c(0,15)) +
  ggtitle("Histograms of Log Population of City Observations Included in the\nVarious Regression Analyses") +
  labs(caption = "Figure 5: This figure makes clear the differences in city size (in terms of population) among the different\nanalyses. It is clear that the listwise deletion of observations with missing values resulted in a sample\nthat almost entirely consisted of larger cities in Trounstine (2016). The distribution of city population\nin Trounstine (2016) more closely resembles the distribution of the original sample with a greater than\n25,000 population cutoff imposed than the original sample. The red lines indicate the median log\npopulation in each sample.") +
  ylab("Count") +
  xlab("Log Population") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.text.x = element_text(size = 7))

@@

```{r over_baqir_t1, echo = FALSE}

@@

kable(imp_over_baqir_1_sum_table_vars,
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for Cities with Populations of 25,000+}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 6716"),
    general = c("This table shows the results of Trounstine's model examining the effect of segregation on overall",
  "per capita direct general expenditure using the multiply imputed dataset for cities with populations above",
  "25,000 people. As in Trounstine (2016), the effect of segregation on public goods spending is large, negative,",
  "and significant while the effect of diversity is small in magnitude and insignificant."),
  footnote_order = c("symbol", "general"), 
           general_title = "Table 3:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

```{r under_baqir_t1, echo = FALSE}

@@

kable(imp_under_baqir_1_sum_table_vars, 
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for Cities with Populations of Less Than 25,000}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 66,403"),
    general = c("This table shows the results of Trounstine's model examining the effect of segregation on overal",
  "per capita direct general expenditure using the multiply imputed dataset for cities with populations below",
  "25,000 people. In contrast to Trounstine (2016) and the results of our analysis with the imputed datasets for",
  "cities over 25,000 people, the effect of segregation on public goods spending is negligible and insignificant,",
  "while the effect of diversity is large, positive, and significant."),
  footnote_order = c("symbol", "general"),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|          As a summary of all of the model results for the two main coefficients of interest, segregation index and diversity, a coefficient plot is presented as Figure XXX, with the models ordered by increasing median log population size of the sample on the x-axis. Looking first at the coefficients for diversity, the two models with low median city populaiton size have significant, positive coefficients, while the models with the larger median city populations have insignificant coefficients. This suggestst that diversity is an important estimator of public goods spending for small, but not necessarily for large, cities. Next, looking at the plot for the segregation index, we see the oppposite effect. For the models where the median city population sizes were small, the coefficient on segregaiton is insignificant. However, for the two models with larger median city populations, the coefficients were significant and negative. This suggests that for small cities, segregation is not a very good estimator of public goods spending, but when cities are larger, increases in a city's segregation index are associated with decreases in public goods spending.

@@

model_summaries %>%
  ggplot(aes(x = median, y = estimate, color = Model)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "black") +
  facet_wrap(~variable) +
  ggtitle("Coefficient Plots for Diversity and White/Non-White Segregation Indices\nAcross the Models")+
  xlab("Models Ordered by Increasing Median Population Size of the Sample") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public goods spending\nvaries with city size. As city populaiton size increases, diversity tends to have less explanatory power for\npublic goods spending while segregation has more. The opposite is true when considering city population\nsize decreasing: segregation becomes less important as diversity becomes increasingly important.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0)) +
  labs(color = "Model Category")

@@

|          Why cities differ in their amount of spending on public goods is an interesting question that is frequently debated in the literature. @trounstine_segregation_2016 posits that segregation may be an important underlying cause of the underprovsion of public goods. Finding evidence that residential segregation by race increases political partisanship, @trounstine_segregation_2016 argues that residential segregation can increase division and make cooperation amongst competing groups difficult, which results in decreased spending on public goods. The findings of @trounstine_segregation_2016 are particularly important because they suggest that segregation, as defined by many homogeneous neighborhoods within a larger diverse geographical area, is the key factor in disinvestment in public goods. This contradicts previous research which has suggested that levels of diversity were instead most important (see for example @baqir_public_1999 or @hopkins_diversity_2009).

|          In this paper, I first worked to provide a check on the analysis presented in @trounstine_segregation_2016 by attempting to replicate the work in R. I was able to successfully replicate all of the main results, with the exception of a few of a few marginal analyses of the complex, multi-level models which, to my knowledge, are currently not supported by available R packages. I was, however, able to successfully replicate these results in Stata.

|          I then tested the models in @trounstine_segregation_2016 by re-running them with additional data that I imputed from missing values in the original datasets. The large number of missing values in both of the datasets used in @trounstine_segregation_2016's main analysis created a concern for potential bias in the original results and for a lack of representativeness due to the large proportion of observations excluded from the regression analysis because of missing data. Using the mice package in R, I created multiply imputed datasets, upon which I re-ran the models and pooled the results to generate the final model coefficients and parameters.

|          The results of the data imputation exercise differed between the two datasets. For the analyses using the smaller, racial polarization dataset, re-running the models with the imputed data resulted in a smaller magnitude of the coefficient on the segregation index, and also made it statistically insignificant in the main model specification. This suggests that the segregation index may not have as strong of a positive association with political polarization as @trounstine_segregation_2016 suggests, but in general, the imputation exercise supportted the main conclusions of Trounstine's analysis.

|          On the other hand, for the analyses using the much larger financial segregation dataset, the results of the models using the multiply imputed data sharply contrasted the results of @trounstine_segregation_2016, suggesting that diversity, and not segregation, was the best explanatory variable for public goods spending across all cities. While these results originally seemed quite contradictory, I was able reconcile this finding by discovering that the results of @trounstine_segregation_2016 seem to be only really representative of larger cities. By dropping all observations in the original data with missing values for variables included in the regressions, it seems that @trounstine_segregation_2016 inadvertently restricted their sample to almost entirely larger cities (above 25,000 residents) whereas the full sample consisted of cities that had smaller populaitons on average. To test the hypothesis that diversity is more important for predicting public goods spending for small cities while segregation is more important for predicting public goods spending for larger cities, I divided the original dataset into two subsets based on the city classificaiton criteria used in @baqir_public_1999 that cities must have a population of at least 25,000 people. I then multiply imputed the missing values for each data subset and re-ran the models. I found that the results presented initial evidence in support of my hypothesis. For the cities below the 25,000 person population cutoff, the coefficient for diversity was positive and significant while the coefficient for the segregation index was insignificant. For cities above the 25,000 person cutoff, the results were very similar to the results found in @trounstine_segregation_2016: the coefficient for diversity was insignificant, but the coefficient for the segregation index was significant and negative.

>>>>

# Introduction

|        There is a large degree of variation in public goods spending across local governments. As a result, many scholars have worked to determine what factors may lead to the underprovision of public goods. Research in the past has associated racial diversity or changes in levels of diversity with the underprovision of public goods [@baqir_public_1999; @hopkins_diversity_2009]. However, @trounstine_segregation_2016 argues that it is racial segregation, not diversity in and of itself, that results in the underprovision of public goods. Trounstine’s analysis consists of two main parts. First, she uses election and demographic data from 25 of America’s largest cities between 1990 and 2010 to run a multilevel mixed-effects linear regression with fixed effects for region and year and with random effects for cities in order to show that political polarization increases with segregation. Trounstine measures segregation with Theil's H Index, which measures the degree to which the diversity of a neighborhood differs from the diversity of an entire city. The second main part of Trounstine’s analysis looks at the ability of the Theil’s H Segregation Index to explain a variety of types of public expenditures by city using a sample of 2,637 cities with 13,742 city-year observations. Using linear regressions with fixed effects for cities and robust standard errors clustered by city, Trounstine finds a significant, negative relationship between segregation and public goods spending that is robust to the inclusion of a variety of relevant controls and an alternative specification in which the number of waterways is used as an instrument for segregation.

@@

|        One concern with @trounstine_segregation_2016 is the exclusion of a large proportion of observations from the regression analysis due to missing data for some values, which substantially constrains the sample size. For example, the regression analysis in the main specification using the the original racial polarization dataset exludes %`r round((obs_rp - 91)/obs_rp * 100, 0)` of the observations, and the regression analysis in the main specification using the original financial segregation dataset excludes %`r round((obs_fin_seg - 13742)/ obs_fin_seg * 100, 0)` of observations. As an extension of @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. Then, I use the multiply imputed datasets to re-estimate the original models, pooling the results to generate final pooled regression coefficients and parameters.

@@

|        In order to understand what segregation is and why it is important as a political concept, it is critical to understand the literature on the effects of diversity at various geographic levels. Studying diversity can be somewhat difficult and confusing because its effects tend to differ by geographic level. On the neighborhood level, a lack of racial diversity has been associated with racial intolerance, resentment, and competition between racial groups [@oliver_paradoxes_2010]. Living within non-diverse or segregated neighborhoods has also been associated with holding negative stereotypes and perceptions about out groups [@eric_oliver_intergroup_2003]. As a result, non-diverese, homogeneous neighborhoods have been associated with increased racial tension and political polarization in comparison to integrated, diverse neighborhoods. However, at the city or metropolitan level, the opposite seems to be true: when considering larger geographic areas, diversity, not homogeneity, is correlated with racial tension, competition, prejudice, lower levels of cooperation, and lower spending on public goods [@oliver_paradoxes_2010; @baqir_public_1999; @hopkins_diversity_2009]. Considering these theories jointly would predict that the geographic areas most ripe for racial tension and competition will be those that are diverse overall, but have many homogeneous, non-diverse neighborhoods, meaning that while people of different races co-exist within a larger metropolitan area, they live separately within their own neighborhoods. This situation corresponds to @trounstine_segregation_2016's definition of a severely segregated city. The differing effects of diversity at the neighborhood versus the larger geographic level provide an argument for considering segregation as opposed to simply diversity in local politics because measures of segregation (such as the Theils's H Index used by @trounstine_segregation_2016) take into account the patterns of racial mixing at both the neighborhood and larger geographic level that, alltogether, are important in predicting whether a locality will be characterized by interracial cooperation or political polarization [@trounstine_segregation_2016; @oliver_paradoxes_2010; @bharathi_public_2018].

@@

|        In an attempt to better deal with the problem of missing data in @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. While there are a variety of different imputation methods that could have been employed, multiple imputation (such as the multiple multivariate imputations generated by mice) is desirable because instead of inputting a single value such as the mean for missing values, it instead uses the distribution of the available data to estimate multiple potential values for the missing values. As a result, multiple imputation helps to account for the uncertainty inherent in the imputation process and allows for the calculation of standard errors around estimators. This allows the researcher to more accurately assess the uncertainty of the analysis in general.

@@

|        In Figure 2, we can similarly observe the trends for the financial segregation dataset. In this dataset, there is a very high proportion of missing values for most of the variables. For example, there are `r format(sum(is.na(fin_imp$H_citytract_NHW_i)), big.mark = ",")` missing values for the segregation index, which is the main independent variable, or %`r round(sum(is.na(fin_imp$H_citytract_NHW_i))/nrow(fin_imp)*100,0)` of the data. In this dataset, there are also many more observations for which values are missing for multiple variables in comparison to the racial polarization dataset, as can be seen in the plot on the right in Figure 2. In total, a mere %`r .0058*100` of observations are complete for all variables included in the main specification for the financial segregation dataset.

@@

imp_1_rp <- mice(rp_impute, method = "cart", maxit = 20)

@@

imp_fs <- mice(fin_imp, method = meth, maxit = 5)

@@

```{r rp_densityplot, echo = FALSE, fig.cap = "This figure presents the density plots of the imputed data (red) overlaid on the density plot for the pre-existing data (blue) for all those variables with missing values in the racial polarization dataset. For the most part, the imputed data seem to fit the pre-existing data relatively well. The only cases in which there is some cause for concern were for some of the variables with very few missing values in the racial polarization dataset (such as the indicators for region). However, since there are so few of these values missing in the actual datasets, this was not a major concern.", fig.pos="H"}

@@

kable(imp_1_rp_sum_table_vars, 
      "latex", 
      caption = "\\textbf{Racial Polarization in Segregated Cities using the Multiply Imputed Data: Racial Divide with Multigroup Segregation Index}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 203"),
    general = c("This table shows the results of Trounstine's first model for the effect of segregation on political polarization",
  "using the multi-group Theil's H Segregation Index and the new, multiply imputed datasets. The sign",
  "of the coefficient on the main variable of interest, the segregation index, is positive as in Trounstine",
  "(2016), but the result has become statisically insignificant with the 95% confidence interval containing",
  "zero. Thus while these results still suggest that segregation may be associated with increased",
  "political polarization, they confer a lesser degree of certainty than Trounstine's original analysis.",
  "In general, the standard errors on the coefficients have decreased slightly."),
           general_title = "Table 1:",
  footnote_order = c("symbol", "general"))%>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|        For the second and third model specifications, which use the two-group Theil's H Segregation Index, the results were similar. The main coefficient on the segregation index decreased in magnitude, although it remained statistically significant. The other coefficients except for the biracial indicator are insignificant, and standard errors have reduced slightly in these specifications compared to the original analysis in Trounstine (2016). As a visual comparison of the model coefficients, I include coefficient plots for the diversity and segregation index variables as Figure 4. From this plot, it is clear that across all of the models, the coefficient on diversity is positive and insignificant, while the coefficient on the segregaiton index is positive and, in most cases, significant.

@@

|        For the financial segregation dataset, I similarly fit each of our 5 imputed datasets to the original models from @trounstine_segregation_2016 and then pooled the results for each. As a result of the imputation process, I was able to include an additional `r format(73119 - 13742, big.mark = ",")` observations in the analysis, increasing the sample size by %`r round((73119 - 13742)/ 13742*100,0)`. While this is an improvement, I was unable to impute all of the missing values because of the large percentage of missing values in the original dataset, which limited the performance of the algorithm, and due to time and computing constraints for this project.

@@

kable(imp_felm1_sum_table_vars, 
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for all Cities}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 73,119"),
    general = c("This table shows the results of Trounstine's first model for the effect of segregation on per capita direct",
  "public expenditure using the two-group Theil's H Segregation Index and the new, multiply imputed datasets.",
  "The results present a stark contrast to Trounstine's original results: the main coefficient on segregation index",
  "has decreased to essentially zero and has become statistically insignificant.  The coefficient on diversity has",
  "increased in magnitude and become statistically significant."),
  footnote_order = c("symbol", "general"),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|        Because these results differed to such a degree from @trounstine_segregation_2016's original findings, I performed some checks on the data included in each analysis. I discovered that @trounstine_segregation_2016 excluded most cities below a certain population size because of the nature of the paper's list-wise deletion of observations containing missing values: as small cities were more likely to be missing data, it seems that they were inadvertently and disproportionately dropped from @trounstine_segregation_2016's regression analysis. Given this development, I decided to test whether or not city population size affected the results. Making use of the population cutoff of at least 25,000 people used to define cities in @baqir_public_1999, I divided the orginal sample into cities having populations above and below 25,000 people. I then again imputed missing values for each and used the multiply imputed datasets in re-runs of Trounstine's original models.

|        Before discussing the results of these two additional analyses, I first illustrate the differences in the city population sizes for the observations used in the various analyses in Figure 5, which displays the population distributions of the samples used in each. The figure makes it clear that the observations included in @trounstine_segregation_2016's original analysis were not representative of the original data sample with the missing values, as most of the cities with lower populations were excluded as a result of the list-wise data deletion. While the dataset generated from the data imputation on the whole sample also has a different distribution for city populaiton, the range of the data seems to be much more representative of the overall data range in comparison to the @trounstine_segregation_2016 sample. In the plot, I also include the distributions of city population size of the multiply imputed datasets above and below the 25,000 population cutoff. For cities above the cutoff, the city population distribution is similar to the distribution in @trounstine_segregation_2016.

@@

imp_fs_pop_overbaqir <- mice(fin_imp_pop_over_baqir, method = meth, maxit = 5)

@@

imp_fs_pop_under_baqir <- mice(fin_imp_pop_under_baqir, method = meth, maxit = 5)

@@

pop_hist_data %>%
  ggplot() +
  geom_histogram(aes(x = log_pop, fill = Analysis), binwidth = .5) +
  geom_vline(aes(xintercept = median_log_pop), color = "red") +
  facet_wrap(~Analysis) +
  coord_cartesian(xlim = c(0,15)) +
  ggtitle("Histograms of Log Population of City Observations Included in the\nVarious Regression Analyses") +
  labs(caption = "Figure 5: This figure makes clear the differences in city size (in terms of population) among the different\nanalyses. It demonstrates how the list-wise deletion of observations with missing values resulted in a\nsample that almost entirely consisted of larger cities in Trounstine (2016). The distribution of city\npopulation in Trounstine (2016) more closely resembles the distribution of the original sample with a\ngreater than 25,000 population cutoff imposed than the distribution of the original sample. The red lines\nindicate the median log population in each sample.") +
  ylab("Count") +
  xlab("Log Population") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.text.x = element_text(size = 7))

@@

```{r over_baqir_t1, echo = FALSE, fig.pos="H"}

@@

kable(imp_over_baqir_1_sum_table_vars,
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for Cities with Populations of 25,000+}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 6,716"),
    general = c("This table shows the results of Trounstine's model examining the effect of segregation on overall per",
  "capita direct general expenditure using the multiply imputed dataset for cities with populations above",
  "25,000 people. As in Trounstine (2016), the effect of segregation on public goods spending is large, negative,",
  "and significant while the effect of diversity is small in magnitude and insignificant."),
  footnote_order = c("symbol", "general"), 
           general_title = "Table 3:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

```{r under_baqir_t1, echo = FALSE, fig.pos="H"}

@@

kable(imp_under_baqir_1_sum_table_vars, 
      "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for Cities with Populations of Less Than 25,000}",
      align = c("l","c","c","c"),
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(symbol = c("N = 66,403"),
    general = c("This table shows the results of Trounstine's model examining the effect of segregation on overal per",
  "capita direct general expenditure using the multiply imputed dataset for cities with populations below",
  "25,000 people. In contrast to Trounstine (2016) and the results of our analysis with the imputed datasets",
  "for cities over 25,000 people, the effect of segregation on public goods spending is negligible and",
  "insignificant, while the effect of diversity is large, positive, and significant."),
  footnote_order = c("symbol", "general"),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE, latex_options = "hold_position")

@@

|        As a summary of all of the model results for the two main coefficients of interest, segregation index and diversity, a coefficient plot is presented as Figure 6, with the models ordered by increasing median log population size of the sample on the x-axis. Looking first at the coefficients for diversity, the two models with low median city population size have significant, positive coefficients, while the models with the larger median city populations have insignificant coefficients. This suggestst that diversity is an important estimator of public goods spending for small, but not necessarily for large, cities. Next, looking at the plot for the segregation index, we see the oppposite effect. For the models where the median city population sizes were small, the coefficient on segregation is insignificant. However, for the two models with larger median city populations, the coefficients are significant and negative. This suggests that for small cities, segregation is not a very good estimator of public goods spending, but when cities are larger, increases in a city's segregation index are associated with decreases in public goods spending.

@@

model_summaries %>%
  ggplot(aes(x = median, y = estimate, color = Model)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "black") +
  facet_wrap(~variable) +
  ggtitle("Coefficient Plots for Diversity and White/Non-White Segregation Indices\nAcross the Models")+
  xlab("Models Ordered by Increasing Median Population Size of the Sample") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public goods spending\nvaries with city size. As city population size increases, diversity tends to have less explanatory power for\npublic goods spending while segregation has more. The opposite is true when moving in the opposite\ndirection: as city population size decreases, segregation becomes less important as diversity becomes\nincreasingly important in explaining public goods spending.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0)) +
  labs(color = "Model Category")

@@

|        Why cities differ in their amount of spending on public goods is an interesting question that is frequently debated in the literature. @trounstine_segregation_2016 posits that segregation may be an important underlying cause of the underprovsion of public goods. Finding evidence that residential segregation by race increases political partisanship, @trounstine_segregation_2016 argues that residential segregation can increase division and make cooperation amongst competing groups difficult, which results in decreased spending on public goods. The findings of @trounstine_segregation_2016 are particularly important because they suggest that segregation, as defined by many homogeneous neighborhoods within a larger, diverse geographical area, is the key factor in disinvestment in public goods. This contradicts previous research which has suggested that levels of diversity were instead most important (see for example @baqir_public_1999 or @hopkins_diversity_2009).

|        In this paper, I first worked to provide a check on the analysis presented in @trounstine_segregation_2016 by attempting to replicate the work in R. I was able to successfully replicate all of the main results, with the exception of the marginal analyses of the complex, multi-level models which, to my knowledge, are currently not supported by available R packages or functions. I was, however, able to successfully replicate these results in Stata.

|        I then tested the models in @trounstine_segregation_2016 by re-running them with additional data that I imputed from missing values in the original datasets. The large number of missing values in both of the datasets used in @trounstine_segregation_2016's main analysis created a concern for potential bias in the original results because observations with missing values were automatically excluded from the regression analysis. Using the mice package in R, I created multiply imputed datasets, upon which I re-ran the models and pooled the results to generate the final model coefficients and parameters.

|        The results of the data imputation exercise differed between the two datasets. For the analyses using the smaller, racial polarization dataset, re-running the models with the imputed data resulted in a smaller magnitude of the coefficient on the segregation index, and also made it statistically insignificant in the main model specification. This suggests that segregation is still positively associated with political polarization, but maybe not quite as strongly as @trounstine_segregation_2016 suggests.

|        On the other hand, for the analyses using the much larger financial segregation dataset, the results of the models using the multiply imputed data sharply contrasted the results of @trounstine_segregation_2016, suggesting that diversity, and not segregation, was the best explanatory variable for public goods spending across all cities. While these results originally seemed quite contradictory to @trounstine_segregation_2016, I was able reconcile this finding by discovering that the results of @trounstine_segregation_2016 seem to be representative only of larger cities. By dropping all observations in the original data with missing values for variables included in the regressions, it seems that @trounstine_segregation_2016 inadvertently restricted their sample to almost entirely larger cities (above 25,000 residents) whereas the full sample consisted of cities that had smaller populations on average. To test the hypothesis that diversity is more important for predicting public goods spending for small cities while segregation is more important for predicting public goods spending for larger cities, I divided the original dataset into two subsets of cities with populations above and below 25,000 people. I then multiply imputed the missing values for each data subset and re-ran the models. I found that the results presented initial evidence in support of my hypothesis. For the cities below the 25,000 person population cutoff, the coefficient for diversity was positive and significant while the coefficient for the segregation index was insignificant. For cities above the 25,000 person cutoff, the results were very similar to the results found in @trounstine_segregation_2016: the coefficient for diversity was insignificant, but the coefficient for the segregation index was significant and negative.
